BivData Setup

### 2020-01-21 doc/Bivalent_Analysis.Rmd
merged in manual whole cell data to the BivData

#load Clean BivData
load(file="data/CleanBivData_1.20.20.RData")

@@

```{r tech.ect, echo=FALSE, eval=FALSE}

nBiv.per.cell <- ggplot(cells.BivData_table, aes(nbivs)) + ggtitle("Number of passing bivalents per cell") + geom_histogram(bins = 30) +facet_wrap( ~ category)
nBiv.per.cell

@@

@@

@@

**Pro Bivalent level**  
- The data are continous and closer to normal/gaussian expectations than the MLH1 counts.

@@

**Pro Mouse level averages**  
- Mouse average is a conserved framework, it would summarize general patterns which fits with the paper.

@@

# Heterochiasmy Predictions

@@

1) SC length (cite)

@@

# SC lengths

```{r show.SC.length}
#compile data
#main figures, show both mouse and bivalent level

@@

```{r justify.mouse.level}

@@

```{r MM.SC.Length}

@@

@@

```{r show.IFD, echo=FALSE}

@@

Double Check the dataset for the mixed models.

@@

@@

```{r CO.positions}
#main figure of density plots,
#MELT_Curated_BivData.PER

MELT_Curated_BivData.PER <- add_subsp(MELT_Curated_BivData.PER)

Musc.female_MELT_curBivData <- MELT_Curated_BivData.PER[ ((MELT_Curated_BivData.PER$sex == "female") & (MELT_Curated_BivData.PER$subsp == "Musc") ), ]

CO.Pos.density_Musc_female <- ggplot(Musc.female_MELT_curBivData, aes(x=PER_Position, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc female \nNormalized Foci Positions")
CO.Pos.density_Musc_female

#Dom female
Dom.female_MELT_curBivData <- MELT_Curated_BivData.PER[ ((MELT_Curated_BivData.PER$sex == "female") & (MELT_Curated_BivData.PER$subsp == "Dom") ), ]

CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER_Position, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
CO.Pos.density_Dom_female

#Musc male
Musc.male_MELT_curBivData <- MELT_Curated_BivData.PER[ ((MELT_Curated_BivData.PER$sex == "male") & (MELT_Curated_BivData.PER$subsp == "Musc") ), ]

CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER_Position, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
CO.Pos.density_Musc_male

#Dom male
Dom.male_MELT_curBivData <- MELT_Curated_BivData.PER[ ((MELT_Curated_BivData.PER$sex == "male") & (MELT_Curated_BivData.PER$subsp == "Dom") ), ]

CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER_Position, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
CO.Pos.density_Dom_male

@@

```{r MM_CO_pos, echo=FALSE}

#use the 1CO
#Use MELT DF
#MELT_Curated_BivData

Pos.CO.avs <- ddply(MELT_Curated_BivData.PER, c("mouse", "Foci_Type"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),

                    mean_PER_pos = as.numeric(format(round(  mean(PER_Position), 3 ), nsmall=3) )
)

@@

#isolate the 1COPos.CO.avs
Pos.CO.avs_1CO <- Pos.CO.avs[Pos.CO.avs$Foci_Type == "PER_Foci_1",]

@@

BivData.CO.PER.POS.MixedModel.HQ <- lme(mean_PER_pos ~ subsp * sex, data=Pos.CO.avs_1CO, random=list(strain=pdDiag(~sex) ) )

as.data.frame( round(  summary(BivData.MixedModel.HQ)$tTable, 6 ) )
#sex is significant
```

@@

@@

#F1.x.F2 <- ggplot(MELT_Curated_BivData.PER, aes(x=F1, fill=F2) ) + geom_density(alpha=.4) + facet_wrap(~hand.foci.count)+ggtitle("Normalized Foci Positions")

>>>>

#load Clean BivData
load(file="data/CleanBivData_1.21.20.RData")

@@

```{r tech.ect, echo=FALSE, eval=TRUE}

nBiv.per.cell <- ggplot(cell.level.BivData_table, aes(nbivs)) + ggtitle("Number of passing bivalents per cell") + geom_histogram(bins = 30) +facet_wrap( ~ category)
nBiv.per.cell

@@

#This isn't the same as curated/passing bivalents

@@

```{r Chrm.props, echo=FALSE}

#barplots for proportions of chrm classes

#consider adding this the the setup block (could have done this to category level table)
mouse.level.BivData_table$per.0CO <- mouse.level.BivData_table$CO0 / mouse.level.BivData_table$nboxes_passed

mouse.level.BivData_table$per.1CO <- mouse.level.BivData_table$CO1 / mouse.level.BivData_table$nboxes_passed

mouse.level.BivData_table$per.2CO <- mouse.level.BivData_table$CO2 / mouse.level.BivData_table$nboxes_passed

mouse.level.BivData_table$per.3CO <- mouse.level.BivData_table$CO3 / mouse.level.BivData_table$nboxes_passed

mouse.level.BivData_table$per.4CO <- mouse.level.BivData_table$CO4 / mouse.level.BivData_table$nboxes_passed

#DF needs to be melted -- so they all in 1 col
#per code, I've done this by hand

#subset DF for ease
skinny.mouse.level <- mouse.level.BivData_table[-c(11:15 ) ]
col.names.list.99 <- colnames(skinny.mouse.level)

#"mouse"         "ncells"        "boxs.IDd"      "nboxes_passed" "Nbiv"          "CO0"           "CO1"       #[8] "CO2"           "CO3"           "CO4"           "per.0CO"       "per.1CO"       "per.2CO"       "per.3CO"      [15] "per.4CO"

#all of the cols to use
MELT_Chrm.Proportion <- melt(skinny.mouse.level[,c(  "mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","CO0","CO1","CO2", "CO3","CO4",

               #leave out the Foci
                "per.0CO", "per.1CO", "per.2CO", "per.3CO", "per.4CO")], 
  id=c("mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","CO0","CO1","CO2", "CO3","CO4"))

#MELT_Chrm.Proportion <- MELT_Chrm.Proportion[(!is.na(MELT_Chrm.Proportion$value)),]

#MELT_Chrm.Proportion <- MELT_Chrm.Proportion[!(is.na(MELT_Chrm.Proportion$value) | MELT_Chrm.Proportion$value==0), ]

MELT_Chrm.Proportion <- add_category(MELT_Chrm.Proportion)
MELT_Chrm.Proportion <- add_sex(MELT_Chrm.Proportion)
MELT_Chrm.Proportion <- add_strain(MELT_Chrm.Proportion)

#then this DF has to be summarized into categories
chrm.class.category <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
                             mean.percent.type = mean(value)
                             #sum is the wrong!
                             #used average percentage, might consider a different summary
)

#ggplot

chrm.class.barplot <- ggplot(data = chrm.class.category, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(~sex, scales = "free")
chrm.class.barplot

the above plot displays the average proportion of each type of chromosome class across mice. The results mostly align with the hand measured results.

Look into ordering the mouse strains to highlight the MSM and PWD differences, renameing the chrm type labels, spacing out the x axis labels or removing some of the other mice.

@@

**Pro Bivalent level**

- The data are continous and closer to normal/gaussian expectations than the MLH1 counts.

@@

**Pro Mouse level averages**

- Mouse average is a conserved framework, it would summarize general patterns which fits with the paper.

@@

## Predictions, Heterochiasmy

@@

1) SC length (cite Lynn)

@@

# SC Lengths

```{r show.SC.length, echo=FALSE}
#compile data
#main figures, show both mouse and bivalent level

@@

TRY ADDING NESTED boxplots for hand.foci.count

@@

Above is the model we are testing evolution of SC lengths.

```{r MM.SC.Length, echo=FALSE}

@@

The results seem to indicate that sex is a significant factor. Consider writing a subsampling approach (randomize / permute a data set of BivData)

```{r justify.mouse.level, echo=FALSE}

#start the permutation code here
#make a randomized BivData DF

#30 cells worth of bivs? 
#30 cells * ( 19 or 20 bivs ) * 70% segment rate = decent sample size for comparing Biv-based patterns?
# I think this would address if the sample sizes are affecting the pattern

#

@@

**I**nter**f**ocal **D**istance, a indication of CO interference.

DO I have anything to say....

How can I highlight the result (I need to verify this first) for greater IFD in high rec male mice?

The raw measures are pretty nosy, this might not be the best way to display these results.

```{r show.IFD, echo=FALSE, warning=FALSE}

@@

$$mouse \ average \ normalized IFD ~=~ subsp * sex + rand(strain* sex) + \varepsilon $$

I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure

@@

sex may be a significant effect for the percent (normalized IFD). Is there a way to measure / account for the SC.length effect?

@@

```{r CO.positions}
#main figure of density plots,

MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain

Musc.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Musc") ), ]

CO.Pos.density_Musc_female <- ggplot(Musc.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc female \nNormalized Foci Positions")
CO.Pos.density_Musc_female

#Dom female
Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]

CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
CO.Pos.density_Dom_female

#Musc male
Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]

CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
CO.Pos.density_Musc_male

#Dom male
Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]

CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
CO.Pos.density_Dom_male

@@

```{r MM_CO_pos, echo=FALSE}
#use the 1CO
#Use MELT DF
#MELT_Curated_BivData

Pos.CO.avs <- ddply(MELT_Curated_BivData, c("mouse", "Foci_Type"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),

                    mean_PER_pos = as.numeric(format(round(  mean(PER.Foci.Pos), 3 ), nsmall=3) )
)

@@

#isolate the 1COPos.CO.avs
Pos.CO.avs_1CO <- Pos.CO.avs[Pos.CO.avs$Foci_Type == "Foci1",]

@@

BivData.CO.PER.POS.lme <- lme(mean_PER_pos ~ subsp * sex, data=Pos.CO.avs_1CO, random=list(strain=pdDiag(~sex) ) )

as.data.frame( round(  summary(BivData.CO.PER.POS.lme)$tTable, 6 ) )
#sex is significant
```

@@

CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]

CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]

@@

F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
F1.x.F2

### 2020-01-21 src/BivData_Setup.Rmd
merged in manual whole cell data to the BivData

@@

#write the curated file
#write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_21.1.20.csv", sep=",",
#            row.names = TRUE)
#
#write.table(OG.cURATE.table.category, "~./MLH1repo/data/CurrentCuratedDataset_cat_21.1.20.csv", sep=",",
#            row.names = TRUE)

@@

@@

@@

`r head(cell.level.BivData_table)`

@@

@@

`r head(mouse.level.BivData_table)`

@@

# Number of whole cells

- list the cells with all bivs measued or which are close to full.

>>>>

whole.cell.data_manual.OG <- read.csv("~./MLH1repo/whole_cell_measures.csv")

@@

#write the curated file
#write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_21.1.20.csv", sep=",",
#            row.names = TRUE)
#
#write.table(OG.cURATE.table.category, "~./MLH1repo/data/CurrentCuratedDataset_cat_21.1.20.csv", sep=",",
#            row.names = TRUE)
```

```{r split.up.DF, echo=FALSE}

@@

# Manual Whole Cell

```{r whole.cell.manual.clean, echo=FALSE}
#compare and clean up whole cell measures

whole.cell.data_manual <- whole.cell.data_manual.OG[whole.cell.data_manual.OG$chrm.class != "XY",]

colnames(whole.cell.data_manual.OG)
#"return5"         "fileName"        "boxNumber"       "centromere"      "SC.length"  #    
 #[6] "rank"            "chrm.class"      "hand.foci.count" "F1"              "F2"    #         
#[11] "F3"              "notes"           "X"               "X.1"            
#remove, X, X.1
whole.cell.data_manual <- whole.cell.data_manual[-c(1, 13:14)]
whole.cell.data_manual$Foci4 <- ""

#rename foci positions
whole.cell.data_manual$Foci1 <- whole.cell.data_manual$F1
whole.cell.data_manual$Foci2 <- whole.cell.data_manual$F2
whole.cell.data_manual$Foci3 <- whole.cell.data_manual$F3

#mouse, strain, sex, category
whole.cell.data_manual <- add_mouse(whole.cell.data_manual)
whole.cell.data_manual <- add_category(whole.cell.data_manual)
whole.cell.data_manual <- add_strain(whole.cell.data_manual)
whole.cell.data_manual <- add_sex(whole.cell.data_manual)
whole.cell.data_manual <- add_subsp(whole.cell.data_manual)

#whole.cell.data_manual <-   whole.cell.data_manual[(!is.na(whole.cell.data_manual$fileName)),]
whole.cell.data_manual <- whole.cell.data_manual[!(is.na(whole.cell.data_manual$fileName) | whole.cell.data_manual$fileName==""), ]

#obj.id
whole.cell.data_manual$Obj.ID <- paste(whole.cell.data_manual$fileName, whole.cell.data_manual$boxNumber, sep = "_")

#add all the extra columns, blah

#percent Foci position

#ADD RANK
for(cellys in as.character(unique(whole.cell.data_manual$fileName))){

 # print(cellys)
  chrm.order = rank(whole.cell.data_manual$SC.length[whole.cell.data_manual$fileName == cellys])
  whole.cell.data_manual$rank[whole.cell.data_manual$fileName == cellys] <- chrm.order  
  }

#IFD 
whole.cell.data_manual <- add_IFD(whole.cell.data_manual)

ToDo

- Go back and clean up the original excel data

- clean the bottom, fill in missing data

- check the order of foci positions for 5aug17_14jul17_LEW_f1_sp1_11_rev_blob4.

@@

as.data.frame( head(cell.level.BivData_table) )

@@

```{r whole.cell_merging, echo=FALSE}

#check if there are fileNames
double.filename <- whole.cell.data_manual[ whole.cell.data_manual$fileName %in% Curated_BivData$fileName, ]

#which cell-levels fileNames have close to 15-17 chrms?

almost_whole_cell <- cell.level.BivData_table[(cell.level.BivData_table$Nbiv < 21),]
almost_whole_cell <- almost_whole_cell[almost_whole_cell$Nbiv > 15,]

#there is some discrpancy with the Nbiv and nbivs columns! ugh why, some have 20 and 10

#get the Biv data 
almost_whole_cell_BivData <- Curated_BivData[ Curated_BivData$fileName %in% almost_whole_cell$fileName, ]

as.data.frame(  table( droplevels( almost_whole_cell_BivData$fileName) ) )
length(unique(almost_whole_cell_BivData$fileName))

#what next?

3 cells have manual measures and BivData measures. test error, need to match them up. There are 145 cells from the BivData set which are close to whole cells.

8jun15_23apr15_PWD_f2_sp1_20_rev, 29nov16_13nov16_MSM_m2_sp1_13_rev, 27oct16_31aug16_MSM_m1_sp1_9_rev

@@

as.data.frame( head(mouse.level.BivData_table) )
#kable didn't work

@@

@@
